---
title: "customer_churn"
author: "Bibek Shah"
date: "`r Sys.Date()`"
output: html_document
---
#Importing libraries 
```{r}
library(ggplot2)
library(dplyr)
library(readr)
library(corrplot)
```
#Loading the dataset
```{r}
data <- read.csv("C:/Users/Public/Documents/sem_4_R/Dataset/OneDrive_2024-09-05/customer churn/Telecom_customer churn.csv")
```
#Data Overview
```{r}
#checking the structure of dataset
str(data)
```
1.showing summary statistics
```{r}
summary(data)
```
```{r}
head(data)
```
2.Viewing the dimension of dataset
```{r}
dim(data)
```
```{r}
# Number of rows
n_rows <- nrow(data)
print(paste("Number of rows:", n_rows))

# Number of columns
n_cols <- ncol(data)
print(paste("Number of columns:", n_cols))

```
3.Viewing column names of the dataset
```{r}
colnames(data)
```
4.Checking the data types of each column
```{r}
sapply(data, class)
```
#Data Manipulation
1.Checking for missing values
```{r}
colSums(is.na(data))
```
#Data Cleaning
2.Handling the missing values and converting categorical variables to factors
```{r}
data <- data %>% mutate_if(is.numeric, ~ ifelse(is.na(.), median(., na.rm = TRUE), .))
data$churn <- as.factor(data$churn)
colSums(is.na(data))
```
#Data Visualization
1.Visualizing churn distribution
```{r}
library(ggplot2)
# Visualize churn distribution
ggplot(data, aes(x = churn)) + 
  geom_bar() + 
  labs(title = "Churn Distribution", x = "Churn", y = "Count")
```
2.Revenue vs Churn graph
```{r}
# Visualize revenue vs churn
ggplot(data, aes(x = totrev, fill = churn)) + 
  geom_histogram(bins = 30) + 
  labs(title = "Total Revenue vs Churn", x = "Total Revenue", y = "Count")
```

3.Visualizing Churn rate by area
```{r}
# churn rate by area
ggplot(data, aes(x = area, fill = churn)) +
  geom_bar(position = "fill") +
  labs(title = "Churn Rate by Area", x = "Area", y = "Proportion") +
  scale_y_continuous(labels = scales::percent) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
4.Visualizing Churn vs. Monthey Revenue
```{r}
# Churn vs Monthly Revenue
ggplot(data, aes(x = totmrc_Mean, fill = churn)) +
  geom_density(alpha = 0.5) +
  labs(title = "Density Plot of Monthly Revenue by Churn", x = "Monthly Revenue", y = "Density") +
  theme_minimal()
```
5.Visualizing Churn vs Mean Outgoing Usage(MOU)
```{r}
# Churn vs. Mean Outgoing Usage
ggplot(data, aes(x = mou_Mean, y = churn, color = churn)) +
  geom_jitter(alpha = 0.5) +
  labs(title = "Churn vs. Mean Outgoing Usage", x = "Mean Outgoing Usage (MOU)", y = "Churn Status") +
  theme_minimal()
```

# Correlation Matrix
```{r}
# Loading the required library for correlation plot
install.packages("corrplot")
library(corrplot)

# Calculating correlation matrix
cor_matrix <- cor(data %>% select_if(is.numeric))

# Plotting the correlation matrix
corrplot(cor_matrix, method = "circle", type = "upper", order = "hclust", 
         tl.col = "black", tl.srt = 45, title = "Correlation Matrix")
```
1.Boxplot of Total Revenue by Churn
```{r}
library(ggplot2)
ggplot(data, aes(x = churn, y = totrev)) +
  geom_boxplot() +
  labs(title = "Total Revenue by Churn Status", x = "Churn Status", y = "Total Revenue") +
  theme_minimal()
```
2.Visualizing churn rate over time by month
```{r}
ggplot(data, aes(x = months, fill = churn)) +
  geom_bar(position = "fill") +
  labs(title = "Churn Rate Over Time", x = "Months", y = "Proportion") +
  scale_y_continuous(labels = scales::percent) +
  theme_minimal()
```
3.Pair plot for a selection of numerical features
```{r}
# Loading the required package for pair plots
install.packages("GGally")
library(GGally)

# Pair plot for a selection of numerical features
ggpairs(data %>% select(totrev, mou_Mean, totmrc_Mean, churn),
        mapping = aes(color = churn)) +
  labs(title = "Pairwise Comparison of Selected Features")
```
4.Stacked bar chart for churn by customer type
```{r}
ggplot(data, aes(x = marital, fill = churn)) +
  geom_bar(position = "fill") +
  labs(title = "Churn Rate by Marital Status", x = "Marital Status", y = "Proportion") +
  scale_y_continuous(labels = scales::percent) +
  theme_minimal()
```
5.Histogram of Monthly Revenue
```{r}
# Histogram of Monthly Revenue
ggplot(data, aes(x = totmrc_Mean, fill = churn)) +
  geom_histogram(bins = 30, alpha = 0.7, position = "identity") +
  labs(title = "Distribution of Monthly Revenue", x = "Monthly Revenue", y = "Count") +
  theme_minimal()
```
6.Violin plot of data usage by churn status
```{r}
ggplot(data, aes(x = churn, y = datovr_Mean, fill = churn)) +
  geom_violin(trim = FALSE) +
  labs(title = "Data Usage Distribution by Churn Status", x = "Churn Status", y = "Data Over Usage Mean") +
  theme_minimal()
```
7.Scatter plot for Total Revenue vs. Mean Outgoing Usage
```{r}
# Scatter plot for Total Revenue vs. Mean Outgoing Usage
ggplot(data, aes(x = totrev, y = mou_Mean, color = churn)) +
  geom_point(alpha = 0.6) +
  labs(title = "Total Revenue vs. Mean Outgoing Usage", x = "Total Revenue", y = "Mean Outgoing Usage (MOU)") +
  theme_minimal() +
  geom_smooth(method = "lm", se = FALSE)
```
8.Boxplot of Outgoing Calls by Churn
```{r}
ggplot(data, aes(x = churn, y = totcalls)) +
  geom_boxplot() +
  labs(title = "Outgoing Calls by Churn Status", x = "Churn Status", y = "Total Calls") +
  theme_minimal()
```
9.Radar Chart of Feature Distributions
```{r}
# Load the required package for radar charts
install.packages("fmsb")
library(fmsb)

# Prepare data for radar chart
feature_data <- data %>%
  group_by(churn) %>%
  summarise(across(c(totmrc_Mean, mou_Mean, totrev, datovr_Mean), mean))

# Normalize the data
feature_data <- rbind(rep(200, 4), rep(0, 4), feature_data)  # Adjust limits as needed
colnames(feature_data) <- c("totmrc_Mean", "mou_Mean", "totrev", "datovr_Mean")
rownames(feature_data) <- c("Max", "Min", "Churned", "Not Churned")

# Radar chart
radarchart(feature_data, axistype = 1, pcol = c("red", "blue"), plwd = 2, plty = 1)
```



